---
title: "Interesting Forecasting Applications with R"
author: "Bahadir Ayan - IE360 - Srping 2020"
---
```{r,echo=FALSE,results="hide"}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

```{r,echo=FALSE,eval=FALSE}
rmarkdown::render("/home/baydogan/Courses/IE360/Guidelines/example_homework_0.Rmd",output_format="pdf_document")
rmarkdown::render("/home/baydogan/Courses/IE360/Guidelines/example_homework_0.Rmd",output_format="html_document")
```


1. [Time series forecasting using ARIMA  models] (https://blogs.oracle.com/datascience/introduction-to-forecasting-with-arima-in-r)

This example is important because it introduces the forecasting methodology using ARIMA models in R. The steps are first load R packages,examine your data, decompose your data,
stationarity, autocorrelations and choosing model order, fitting an ARIMA Model and lastly evaluate and iterate. Example on Oracle explains these 7 steps very briefly with examples.
That example very good for introductory data.

2.[Time Series Analysis Using ARIMA Model in R](https://datascienceplus.com/time-series-analysis-using-arima-model-in-r/)

This example is explaining methods for time series anaylsis.After that, exaple is explaining ARIMA Modeling by applications.
Also that source supported with graphs and R codes. Lastly, source also contains information about model fitting, forecasting with  ARIMA models.

3.[ARIMA modeling with R](https://www.datacamp.com/community/blog/arima-modeling-with-r)

This example is also education tool for the ARIMA modeling with R. As a result, source explains: why ARIMA modeling with R is important, and it's application areas.

4.[ARIMA models and Intervention Analysis](https://www.r-bloggers.com/arima-models-and-intervention-analysis/)

This example is a complex one. It contains 6 different ARIMA models. Explaining methods very detailed and supported with graphs and R codes.

5.[ARIMA model for forecasting-Example in R](http://ftp.uni-bayreuth.de/math/statlib/general/tsa2/Rissues.htm)

This example explains the auto regression and moving average processes. Then, explains the ARIMA model for the time series.
After that, explains the building ARIMA model and gives example in R. This example also supported with graphs and R Codes.
